Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-07-17 Paper/paper.tex
Update

@@

%% JFH Let's make all these conversion factors and codes to do K-correction
%% publicly avaialble in an appendix.

For the $z<2.2$ sample we corrected the SDSS $i$ and 2SLAQ $g$ band
magnitudes for host galaxy contamination following
\citet{2009MNRAS.392...19C}. Considering the different magnitude
limits of 2SLAQ and SDSS, the modelled host galaxy contamination is
small for $z>0.5$ quasars ($<0.1$\,mag in $g$, $<0.2$\,mag in $i$),
and is negligible at $z>0.8$.  In case the band defining the magnitude
limit of the survey undesirably overlaps with the Ly$\alpha$ forest
\citep{2010ApJ...710.1498G, 2011ApJ...728L..26G, 2013ApJ...768..105M}
we adopted their respective bandpass corrections to $M_{1450}$. In
particular, for the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G}
sample we recomputed $M_{1450}$ from the $R$ band photometry to be
consistent with the selection function defined in $R$, and to avoid
uncertainties in their spectrophotometry due to incomplete spectral
coverage. Since the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G}
$R$ band traces the rest frame UV, we assumed negligible host galaxy
contamination for their faint quasars. For the remaining high-redshift
surveys reporting $M_{1450}$ obtained by various methods
\citep{2010AJ....139..906W, 2011Natur.474..616M, 2015ApJ...798...28K,
  2015ApJ...801L..11V, 2016ApJ...829...33Y, 2016ApJ...833..222J,
  2018Natur.553..473B} we did not re-compute $M_{1450}$, but applied
appropriate shifts to correct to our adopted cosmology.

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$
\citep{2013ApJ...773...14R}, and we adopted
%% JFH The overline looks weird here. I would suggest taht you rather
%% use angle brackets instead. It is not jiving with the equation above
%% and is very hard to read. 
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$.
We obtained a rough estimate of the \citet{2015AA...578A..83G}
selection function by comparing the corrected and observed QLFs,
i.e.\ taking $f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$
(see their Table~3).  Finally, for the three $z>6.5$ quasars we
assumed a selection function of unity in a range of $z$ and $M_{1450}$
estimated by the respective survey teams (private communication).

@@

%% GW: Maybe say what we mean by one-sigma uncertainty: This is a credible interval from the posterior distribution. Do we use the highest posterior density interval or the equal-tailed interval?
\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Homogenised quasar luminosity functions at rest-frame wavelength
    $\lambda=1450$\,\AA\ in redshift bins from $z=0.1$ to $6.5$.
    The symbols show our inferred binned luminosity functions from various
    data sets:
    \citet[red]{2010AJ....139.2360S},
    \citet[green]{2009MNRAS.392...19C},
    \citet[dark blue]{2013ApJ...773...14R},
    \citet[light blue]{2011ApJ...728L..26G},
    \citet[yellow]{2016ApJ...829...33Y},
    \citet[DR7 brown]{2013ApJ...768..105M},
    \citet[Stripe 82 teal]{2013ApJ...768..105M},
    \citet[pink]{2016ApJ...833..222J},
    \citet[orange]{2010AJ....139..906W}, and
    \citet[grey]{2015ApJ...798...28K}.
    Open circles in corresponding colours indicate magnitude bins
    excluded due to incompleteness in the respective data sets.  The
    number of AGN before this selection is shown in parentheses; the
    selected number of AGN is shown outside parentheses.  In each
    redshift bin, the black curve shows our fiducial double power law
    model fit, which is represented by the median of the posterior
    probability distribution function.  The grey shaded area shows the
    one-sigma (68.26\%) uncertainty.  See Sections~\ref{sec:binnedlf}
    and \ref{sec:bins} for further details.}
  \label{fig:mosaic}
\end{figure*}

@@

%%GW: Added a note on discarded redshift bins
%%GW: Check on final Nqso needed! The number at z~5 is inconsistent with Fig. 3
\begin{table*}
  % bins_tabulate.py; Nqso added by hand
  \caption{\textbf{Check on final Nqso needed! The number at $z\sim5$ is inconsistent with Fig. 3. We could also have two columns for the number of quasars: before and after magnitude restrictions.}
    Posterior median double power law luminosity function parameters
    and their $1\sigma$ (68.26\%) statistical errors in
    various redshift bins shown in Figure~\ref{fig:evoln}.
    The luminosity function parameters
    $\phi_*$, $M_*$, $\alpha$, and $\beta$ are defined in
    Equation~(\ref{eqn:dpl}), with $\beta$ denoting the faint-end
    slope.  Quasars in each bin have redshifts $z_\mathrm{min}\leq z <
    z_\mathrm{max}$ with a sample mean $\langle z\rangle$.
    The number of quasars in each bin is given by $N_\mathrm{QSO}$.
    %The corresponding luminosity
    %functions are shown in Figure~\ref{fig:mosaic}.  See
    %Section~\ref{sec:bins} for further details.
    }
  \label{tab:bins}
  \begin{tabular}{lccr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{QSO}$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\

    &
    &
    &
    &
    \multicolumn{1}{c}{cMpc$^{-3}$mag$^{-1}$)} &
    &
    & \\
    \hline
    0.31$^a$ & 0.10 & 0.40 & 3632 & -5.62^{+0.09}_{-0.11} & -21.37^{+0.21}_{-0.23} & -2.94^{+0.09}_{-0.10} & -1.22^{+0.18}_{-0.16} \\
    0.50$^a$ & 0.40 & 0.60 & 4686 & -6.13^{+0.04}_{-0.04} & -23.00^{+0.08}_{-0.07} & -3.38^{+0.08}_{-0.08} & -1.36^{+0.04}_{-0.04} \\
    0.72 & 0.60 & 0.80 & 4684 & -6.42^{+0.09}_{-0.08} & -24.03^{+0.14}_{-0.13} & -3.56^{+0.12}_{-0.13} & -1.75^{+0.04}_{-0.04} \\
    0.91 & 0.80 & 1.00 & 5248 & -6.39^{+0.08}_{-0.08} & -24.50^{+0.12}_{-0.12} & -3.69^{+0.11}_{-0.11} & -1.73^{+0.06}_{-0.05} \\
    1.10 & 1.00 & 1.20 & 6566 & -6.54^{+0.03}_{-0.03} & -25.15^{+0.04}_{-0.04} & -4.18^{+0.08}_{-0.08} & -1.73^{+0.02}_{-0.02} \\
    1.30 & 1.20 & 1.40 & 7132 & -6.42^{+0.04}_{-0.04} & -25.29^{+0.06}_{-0.05} & -3.97^{+0.08}_{-0.08} & -1.74^{+0.03}_{-0.03} \\
    1.50 & 1.40 & 1.60 & 7771 & -6.51^{+0.03}_{-0.03} & -25.67^{+0.04}_{-0.04} & -4.28^{+0.08}_{-0.08} & -1.76^{+0.02}_{-0.02} \\
    1.71 & 1.60 & 1.80 & 7421 & -6.29^{+0.03}_{-0.03} & -25.56^{+0.05}_{-0.05} & -3.96^{+0.07}_{-0.07} & -1.60^{+0.03}_{-0.03} \\
    1.98 & 1.80 & 2.20 & 10876 & -6.72^{+0.03}_{-0.03} & -26.26^{+0.04}_{-0.04} & -4.21^{+0.07}_{-0.07} & -1.87^{+0.02}_{-0.02} \\
    2.30$^a$ & 2.20 & 2.40 & 8419 & -6.16^{+0.07}_{-0.06} & -25.49^{+0.12}_{-0.11} & -3.34^{+0.12}_{-0.12} & -1.61^{+0.04}_{-0.04} \\
    2.45$^a$ & 2.40 & 2.50 & 3403 & -6.40^{+0.08}_{-0.07} & -25.86^{+0.14}_{-0.12} & -3.61^{+0.21}_{-0.22} & -1.60^{+0.05}_{-0.04} \\
    2.55$^a$ & 2.50 & 2.60 & 2640 & -6.15^{+0.08}_{-0.08} & -25.34^{+0.17}_{-0.16} & -3.32^{+0.15}_{-0.16} & -1.39^{+0.08}_{-0.07} \\
    2.65$^a$ & 2.60 & 2.70 & 1883 & -5.98^{+0.05}_{-0.06} & -25.15^{+0.13}_{-0.14} & -3.12^{+0.12}_{-0.12} & -1.05^{+0.09}_{-0.08} \\
    2.75$^a$ & 2.70 & 2.80 & 1135 & -6.29^{+0.08}_{-0.07} & -25.94^{+0.15}_{-0.13} & -3.78^{+0.27}_{-0.30} & -1.34^{+0.07}_{-0.06} \\
    2.85$^a$ & 2.80 & 2.90 & 1069 & -6.45^{+0.11}_{-0.10} & -26.22^{+0.21}_{-0.18} & -3.61^{+0.35}_{-0.41} & -1.46^{+0.08}_{-0.07} \\
    2.95$^a$ & 2.90 & 3.00 & 1104 & -6.76^{+0.07}_{-0.06} & -26.53^{+0.11}_{-0.09} & -5.05^{+0.59}_{-0.63} & -1.71^{+0.05}_{-0.04} \\
    3.05$^a$ & 3.00 & 3.10 & 1127 & -6.77^{+0.08}_{-0.07} & -26.48^{+0.12}_{-0.11} & -4.68^{+0.44}_{-0.52} & -1.70^{+0.06}_{-0.05} \\
    3.15$^a$ & 3.10 & 3.20 & 1041 & -7.26^{+0.16}_{-0.11} & -27.11^{+0.22}_{-0.17} & -4.38^{+0.76}_{-1.30} & -1.96^{+0.07}_{-0.05} \\
    3.25$^a$ & 3.20 & 3.30 & 815 & -7.34^{+0.13}_{-0.13} & -27.20^{+0.20}_{-0.22} & -4.39^{+0.69}_{-0.79} & -1.94^{+0.06}_{-0.05} \\
    3.34$^a$ & 3.30 & 3.40 & 510 & -7.54^{+0.21}_{-0.21} & -27.39^{+0.28}_{-0.37} & -4.67^{+1.38}_{-1.44} & -2.08^{+0.09}_{-0.07} \\
    3.44$^a$ & 3.40 & 3.50 & 155 & -6.76^{+0.19}_{-0.20} & -26.62^{+0.40}_{-0.36} & -3.67^{+0.62}_{-0.83} & -1.24^{+0.28}_{-0.23} \\
    3.88 & 3.70 & 4.10 & 1204 & -7.91^{+0.11}_{-0.11} & -27.26^{+0.13}_{-0.12} & -4.83^{+0.33}_{-0.38} & -2.07^{+0.10}_{-0.09} \\
    4.35 & 4.10 & 4.70 & 603 & -8.32^{+0.29}_{-0.25} & -27.37^{+0.37}_{-0.30} & -4.20^{+0.41}_{-0.50} & -2.21^{+0.15}_{-0.13} \\
    4.92 & 4.70 & 5.50 & 266 & -9.02^{+0.30}_{-0.21} & -27.89^{+0.37}_{-0.26} & -4.58^{+0.71}_{-0.80} & -2.30^{+0.11}_{-0.08} \\
    6.00 & 5.50 & 6.50 & 66 & -10.66^{+0.74}_{-1.11} & -29.21^{+1.08}_{-1.89} & -5.05^{+0.76}_{-1.18} & -2.41^{+0.10}_{-0.08} \\
    \hline
  \end{tabular}\\
  \begin{minipage}{13.0cm}
    \textsuperscript{$a$}{Redshift bin not considered in the joint QLF fit due to systematic errors (open circles in Figure~\ref{fig:mosaic}).}
  \end{minipage}
\end{table*}

@@

\item
  In Model 1 we assume that the functions $F_0$, $F_1$ and $F_2$ from
  Equation~\eqref{eqn:global} are Chebyshev polynomials in
  $\left(1+z\right)$, written as
  \begin{equation}
    F_i\left(1+z\right)=\sum_{j=0}^{n_i}c_{i,j}T_j\left(1+z\right)
    \label{eqn:cbs}
  \end{equation}
  for $i\in\{0,1,2\}$, where $c_{i,j}$ are the parameters from
  Equation~\eqref{eqn:global} and $T_j\left(1+z\right)$ are Chebyshev
  polynomials of the first kind. We try successively higher orders of
  Chebyshev polynomials in order to arrive at a good fit with the
  data.  As we discuss below, we find that $\phi_*$, $M_*$ and
  $\alpha$ prefer quadratic, cubic, and linear evolutions in
  $\left(1+z\right)$, respectively.  For the faint-end slope $\beta$
  we adopt a double power law to account for a possible break at
  $z\sim 3$ that is currently not covered with credible data.  Thus we
  write
  \begin{equation}
    F_3\left(1+z\right)=c_{3,0}+\frac{c_{3,1}}{10^{c_{3,3}\zeta}+10^{c_{3,4}\zeta}},
    \label{eqn:beta}
  \end{equation}
  where
  \begin{equation}
    \zeta = \log_{10}\left(\frac{1+z}{1+c_{3,2}}\right),
  \end{equation}
  thus resulting in a five-parameter model with parameters $c_{3,i}$.
  The parameters $c_{3,3}$ and $c_{3,4}$ thus determine the low and
  high redshift slopes of this evolution, with a break at redshift
  $c_{3,2}$.  This is similar to the model of
  \citet{2007ApJ...654..731H}, who also favoured a broken power law
  model for the evolution of the faint-end slope of the bolometric
  luminosity function of quasars.  Model 1 thus has 14 parameters.
  Excluding those deemed to be dominated by systematic errors, as
  discussed in the previous section, all of the remaining AGN from
  Table~\ref{tab:samples} are included while fitting this model.
  \textbf{GW agrees with JFH that this is too short and
    cryptic. Table~\ref{tab:samples} refers to all samples, but we
    exclude some samples, some redshift ranges, and some magnitude
    ranges. I've put a clarification in the first sentence of the
    section, but we might elaborate more.}
  %% JFH where was this discussed?? Nowhere have I seen explicitly stated in the
  %% text which luminosity/redshift ranges you are excluding from the analysis in performing
  %% these fits because you think they are systematics dominated. This just needs to be clearly stated.
  %% Add this to the end of Section 3.2 after you discuss the systematics. State what is excluded and why.
  %% It is perfectly fine if you do this by eye, just state that.

>>>>

For the $z<2.2$ sample we corrected the SDSS $i$ and 2SLAQ $g$ band
magnitudes for host galaxy contamination following
\citet{2009MNRAS.392...19C}. Considering the different magnitude
limits of 2SLAQ and SDSS, the modelled host galaxy contamination is
small for $z>0.5$ quasars ($<0.1$\,mag in $g$, $<0.2$\,mag in $i$),
and is negligible at $z>0.8$.  In case the band defining the magnitude
limit of the survey undesirably overlaps with the Ly$\alpha$ forest
\citep{2010ApJ...710.1498G, 2011ApJ...728L..26G, 2013ApJ...768..105M}
we adopted their respective bandpass corrections to $M_{1450}$. In
particular, for the $z\sim 4$ sample of \citet{2010ApJ...710.1498G,
  2011ApJ...728L..26G} we recomputed $M_{1450}$ from the $R$ band
photometry to be consistent with the selection function defined in
$R$, and to avoid uncertainties in their spectrophotometry due to
incomplete spectral coverage. Since the \citet{2010ApJ...710.1498G,
  2011ApJ...728L..26G} $R$ band traces the rest frame UV, we assumed
negligible host galaxy contamination for their faint quasars. For the
remaining high-redshift surveys reporting $M_{1450}$ obtained by
various methods \citep{2010AJ....139..906W, 2011Natur.474..616M,
  2015ApJ...798...28K, 2015ApJ...801L..11V, 2016ApJ...829...33Y,
  2016ApJ...833..222J, 2018Natur.553..473B} we did not re-compute
$M_{1450}$, but applied appropriate shifts to correct to our adopted
cosmology.

@@

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample \citep{2013ApJ...773...14R} contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$, and we adopted
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$.
We obtained a rough estimate of the \citet{2015AA...578A..83G}
selection function by comparing the corrected and observed QLFs,
i.e.\ taking $f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$
(see their Table~3).  Finally, for the three $z>6.5$ quasars we
assumed a selection function of unity in a range of $z$ and $M_{1450}$
estimated by the respective survey teams (private communication).

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Homogenised quasar luminosity functions at rest-frame
    wavelength $\lambda=1450$\,\AA\ in redshift bins from $z=0.1$ to
    $6.5$.  The symbols show our inferred binned luminosity functions
    from various data sets: \citet[red]{2010AJ....139.2360S},
    \citet[green]{2009MNRAS.392...19C}, \citet[dark
      blue]{2013ApJ...773...14R}, \citet[light
      blue]{2011ApJ...728L..26G}, \citet[yellow]{2016ApJ...829...33Y},
    \citet[DR7 brown]{2013ApJ...768..105M}, \citet[Stripe 82
      teal]{2013ApJ...768..105M}, \citet[pink]{2016ApJ...833..222J},
    \citet[orange]{2010AJ....139..906W}, and
    \citet[grey]{2015ApJ...798...28K}.  Open circles in corresponding
    colours indicate magnitude bins excluded due to incompleteness in
    the respective data sets.  The number of AGN before this selection
    is shown in parentheses; the selected number of AGN is shown
    outside parentheses.  In each redshift bin, the black curve shows
    our fiducial double power law model fit, which is represented by
    the median of the posterior probability distribution function.
    The grey shaded area shows the one-sigma uncertainty (68.26\%
    equal-tailed credible interval).  See Sections~\ref{sec:binnedlf}
    and \ref{sec:bins} for further details.}
  \label{fig:mosaic}
\end{figure*}

@@

\begin{table*}
  % bins_tabulate.py; Nqso and Nqso' added by hand
  \caption{Posterior median double power law luminosity function
    parameters and their $1\sigma$ (68.26\%) statistical errors in
    various redshift bins shown in Figure~\ref{fig:evoln}.  The
    luminosity function parameters $\phi_*$, $M_*$, $\alpha$, and
    $\beta$ are defined in Equation~(\ref{eqn:dpl}), with $\beta$
    denoting the faint-end slope.  Quasars in each bin have redshifts
    $z_\mathrm{min}\leq z < z_\mathrm{max}$ with a sample mean
    $\langle z\rangle$.  The total number of QSOs in each bin is given
    by $N_\mathrm{QSO}^\prime$.  The selected number of QSOs in each
    bin, after excluding faint QSOs due to excess incompleteness is
    given by $N_\mathrm{QSO}$.}
  \label{tab:bins}
  \begin{tabular}{lccrr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{QSO}$ &
    $N_\mathrm{QSO}^\prime$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\

    &
    &
    &
    &
    &
    \multicolumn{1}{c}{cMpc$^{-3}$mag$^{-1}$)} &
    &
    & \\
    \hline
    0.31$^a$ & 0.10 & 0.40 & 3520 & 3632 & -5.62^{+0.09}_{-0.11} & -21.37^{+0.21}_{-0.23} & -2.94^{+0.09}_{-0.10} & -1.22^{+0.18}_{-0.16} \\
    0.50$^a$ & 0.40 & 0.60 & 4555 & 4686 & -6.13^{+0.04}_{-0.04} & -23.00^{+0.08}_{-0.07} & -3.38^{+0.08}_{-0.08} & -1.36^{+0.04}_{-0.04} \\
    0.72 & 0.60 & 0.80 & 3308 & 4684 & -6.42^{+0.09}_{-0.08} & -24.03^{+0.14}_{-0.13} & -3.56^{+0.12}_{-0.13} & -1.75^{+0.04}_{-0.04} \\
    0.91 & 0.80 & 1.00 & 3922 & 5248 & -6.39^{+0.08}_{-0.08} & -24.50^{+0.12}_{-0.12} & -3.69^{+0.11}_{-0.11} & -1.73^{+0.06}_{-0.05} \\
    1.10 & 1.00 & 1.20 & 6367 & 6566 & -6.54^{+0.03}_{-0.03} & -25.15^{+0.04}_{-0.04} & -4.18^{+0.08}_{-0.08} & -1.73^{+0.02}_{-0.02} \\
    1.30 & 1.20 & 1.40 & 6479 & 7132 & -6.42^{+0.04}_{-0.04} & -25.29^{+0.06}_{-0.05} & -3.97^{+0.08}_{-0.08} & -1.74^{+0.03}_{-0.03} \\
    1.50 & 1.40 & 1.60 & 7539 & 7771 & -6.51^{+0.03}_{-0.03} & -25.67^{+0.04}_{-0.04} & -4.28^{+0.08}_{-0.08} & -1.76^{+0.02}_{-0.02} \\
    1.71 & 1.60 & 1.80 & 6788 & 7421 & -6.29^{+0.03}_{-0.03} & -25.56^{+0.05}_{-0.05} & -3.96^{+0.07}_{-0.07} & -1.60^{+0.03}_{-0.03} \\
    1.98 & 1.80 & 2.20 & 10478 & 10876 & -6.72^{+0.03}_{-0.03} & -26.26^{+0.04}_{-0.04} & -4.21^{+0.07}_{-0.07} & -1.87^{+0.02}_{-0.02} \\
    2.30$^a$ & 2.20 & 2.40 & 8419 & 8419 & -6.16^{+0.07}_{-0.06} & -25.49^{+0.12}_{-0.11} & -3.34^{+0.12}_{-0.12} & -1.61^{+0.04}_{-0.04} \\
    2.45$^a$ & 2.40 & 2.50 & 3403 & 3403 & -6.40^{+0.08}_{-0.07} & -25.86^{+0.14}_{-0.12} & -3.61^{+0.21}_{-0.22} & -1.60^{+0.05}_{-0.04} \\
    2.55$^a$ & 2.50 & 2.60 & 2640 & 2640 & -6.15^{+0.08}_{-0.08} & -25.34^{+0.17}_{-0.16} & -3.32^{+0.15}_{-0.16} & -1.39^{+0.08}_{-0.07} \\
    2.65$^a$ & 2.60 & 2.70 & 1883 & 1883 & -5.98^{+0.05}_{-0.06} & -25.15^{+0.13}_{-0.14} & -3.12^{+0.12}_{-0.12} & -1.05^{+0.09}_{-0.08} \\
    2.75$^a$ & 2.70 & 2.80 & 1135 & 1135 & -6.29^{+0.08}_{-0.07} & -25.94^{+0.15}_{-0.13} & -3.78^{+0.27}_{-0.30} & -1.34^{+0.07}_{-0.06} \\
    2.85$^a$ & 2.80 & 2.90 & 1069 & 1069 & -6.45^{+0.11}_{-0.10} & -26.22^{+0.21}_{-0.18} & -3.61^{+0.35}_{-0.41} & -1.46^{+0.08}_{-0.07} \\
    2.95$^a$ & 2.90 & 3.00 & 1104 & 1104 & -6.76^{+0.07}_{-0.06} & -26.53^{+0.11}_{-0.09} & -5.05^{+0.59}_{-0.63} & -1.71^{+0.05}_{-0.04} \\
    3.05$^a$ & 3.00 & 3.10 & 1127 & 1127 & -6.77^{+0.08}_{-0.07} & -26.48^{+0.12}_{-0.11} & -4.68^{+0.44}_{-0.52} & -1.70^{+0.06}_{-0.05} \\
    3.15$^a$ & 3.10 & 3.20 & 1041 & 1041 & -7.26^{+0.16}_{-0.11} & -27.11^{+0.22}_{-0.17} & -4.38^{+0.76}_{-1.30} & -1.96^{+0.07}_{-0.05} \\
    3.25$^a$ & 3.20 & 3.30 & 815 & 815 & -7.34^{+0.13}_{-0.13} & -27.20^{+0.20}_{-0.22} & -4.39^{+0.69}_{-0.79} & -1.94^{+0.06}_{-0.05} \\
    3.34$^a$ & 3.30 & 3.40 & 510 & 510 & -7.54^{+0.21}_{-0.21} & -27.39^{+0.28}_{-0.37} & -4.67^{+1.38}_{-1.44} & -2.08^{+0.09}_{-0.07} \\
    3.44$^a$ & 3.40 & 3.50 & 155 & 155 & -6.76^{+0.19}_{-0.20} & -26.62^{+0.40}_{-0.36} & -3.67^{+0.62}_{-0.83} & -1.24^{+0.28}_{-0.23} \\
    3.88 & 3.70 & 4.10 & 628 & 1204 & -7.91^{+0.11}_{-0.11} & -27.26^{+0.13}_{-0.12} & -4.83^{+0.33}_{-0.38} & -2.07^{+0.10}_{-0.09} \\
    4.35 & 4.10 & 4.70 & 442 & 603 & -8.32^{+0.29}_{-0.25} & -27.37^{+0.37}_{-0.30} & -4.20^{+0.41}_{-0.50} & -2.21^{+0.15}_{-0.13} \\
    4.92 & 4.70 & 5.50 & 263 & 263 & -9.02^{+0.30}_{-0.21} & -27.89^{+0.37}_{-0.26} & -4.58^{+0.71}_{-0.80} & -2.30^{+0.11}_{-0.08} \\
    6.00 & 5.50 & 6.50 & 66 & 66 & -10.66^{+0.74}_{-1.11} & -29.21^{+1.08}_{-1.89} & -5.05^{+0.76}_{-1.18} & -2.41^{+0.10}_{-0.08} \\
    \hline
  \end{tabular}\\
  \begin{minipage}{13.0cm}
    \textsuperscript{$a$}{Redshift bin not considered in the joint QLF fit due to systematic errors (open circles in Figure~\ref{fig:mosaic}).}
  \end{minipage}
\end{table*}

@@

\item
  In Model 1 we assume that the functions $F_0$, $F_1$ and $F_2$ from
  Equation~\eqref{eqn:global} are Chebyshev polynomials in
  $\left(1+z\right)$, written as
  \begin{equation}
    F_i\left(1+z\right)=\sum_{j=0}^{n_i}c_{i,j}T_j\left(1+z\right)
    \label{eqn:cbs}
  \end{equation}
  for $i\in\{0,1,2\}$, where $c_{i,j}$ are the parameters from
  Equation~\eqref{eqn:global} and $T_j\left(1+z\right)$ are Chebyshev
  polynomials of the first kind. We try successively higher orders of
  Chebyshev polynomials in order to arrive at a good fit with the
  data.  As we discuss below, we find that $\phi_*$, $M_*$ and
  $\alpha$ prefer quadratic, cubic, and linear evolutions in
  $\left(1+z\right)$, respectively.  For the faint-end slope $\beta$
  we adopt a double power law to account for a possible break at
  $z\sim 3$ that is currently not covered with credible data.  Thus we
  write
  \begin{equation}
    F_3\left(1+z\right)=c_{3,0}+\frac{c_{3,1}}{10^{c_{3,3}\zeta}+10^{c_{3,4}\zeta}},
    \label{eqn:beta}
  \end{equation}
  where
  \begin{equation}
    \zeta = \log_{10}\left(\frac{1+z}{1+c_{3,2}}\right),
  \end{equation}
  thus resulting in a five-parameter model with parameters $c_{3,i}$.
  The parameters $c_{3,3}$ and $c_{3,4}$ thus determine the low and
  high redshift slopes of this evolution, with a break at redshift
  $c_{3,2}$.  This is similar to the model of
  \citet{2007ApJ...654..731H}, who also favoured a broken power law
  model for the evolution of the faint-end slope of the bolometric
  luminosity function of quasars.  Model 1 thus has 14 parameters.
  Excluding those deemed to be dominated by systematic errors, as
  discussed in the previous section, all of the remaining AGN from
  Table~\ref{tab:samples} are included while fitting this model.

\item Model 2 parameterises the luminosity function evolution in the
  same way as Model 1, so that the faint-end slope evolution is
  described by a double power law while the evolution of the other
  parameters $\phi_*$, $M_*$ and $\alpha$ is described by,
  respectively, quadratic, cubic, and linear polynomials in $(1+z)$.
  The total number of parameters is 14.  However, while fitting this
  model, we exclude the samples from \citet{2015AA...578A..83G},
  \citet{2011Natur.474..616M}, \citet{2015ApJ...801L..11V}, and
  \citet{2018Natur.553..473B} from the analysis (samples 17--20 from
  Table~\ref{tab:samples}).  As discussed in
  Section~\ref{sect:samplesel}, these samples have approximate
  selection functions.  Removing them allows us to understand the
  effect this has on the favoured evolution model.

\item In Model 3, we again exclude samples from
  \citet{2015AA...578A..83G}, \citet{2011Natur.474..616M},
  \citet{2015ApJ...801L..11V}, and \citet{2018Natur.553..473B} from
  the analysis (samples 17--20 from Table~\ref{tab:samples}).  We also
  continue to describe the evolution of $\phi_*$, $M_*$ and $\alpha$
  by quadratic, cubic, and linear polynomials in $(1+z)$,
  respectively.  But in this model, the evolution of the faint-end
  slope $\beta$ is also assumed to be linear in $(1+z)$.  This model
  thus has just 11 parameters.
\end{itemize}
